\section{Discussion}
\label{sec:discussion}

\subsection{Tractable Hamiltonian Classes}
\label{subsec:tractable}

The eigenbasis requirement
(Section~\ref{subsec:eigenbasis}) restricts stereographic block
encoding to structured Hamiltonians.  The three classes identified
in that section---computational-basis diagonal operators,
Hamiltonians with known diagonalization, and QPE-assisted
encoding---cover a range of physically relevant systems.

Diagonal operators arise naturally in quantum chemistry
(diagonal Coulomb integrals), combinatorial optimization
(QUBO/Ising cost functions), and classical simulation.  For
these, stereographic block encoding provides a concrete advantage:
the encoding depends only on individual eigenvalues, not on the
global norm, and the single-ancilla construction is simpler than
standard block encoding.

Hamiltonians with known diagonalization include integrable spin
chains solvable by the Bethe ansatz~\cite{Bethe1931}, free-fermion
models diagonalizable by the Jordan--Wigner
transformation~\cite{JordanWigner1928}, and small systems where
exact diagonalization is tractable.  The Heisenberg model
(Section~\ref{subsec:heisenberg}) is a representative example:
the change of basis is a Givens rotation with $\mathcal{O}(1)$
gate complexity.

For Bethe-ansatz-solvable models, the eigenstates are
parameterized by rapidities satisfying algebraic equations,
and the challenge is converting these solutions into
quantum circuits.  The XXX and XXZ spin chains have
eigenstates expressible as coordinate Bethe wave functions,
which can in principle be prepared by circuits of depth
polynomial in the system size~$n$ and the number of
magnons~$M$.  For the two-magnon sector ($M=2$), the
wave function is a sum of $\binom{n}{2}$ computational
basis states with known amplitudes, and state preparation
via amplitude encoding has gate complexity
$\mathcal{O}(n^2)$.  Whether the full diagonalization
$V$ (across all magnon sectors simultaneously) can be
implemented efficiently is an open question that connects
to the broader program of quantum simulation of integrable
models.

Free-fermion systems (e.g., the transverse-field Ising
model) provide a more immediately tractable class: the
Jordan--Wigner transformation maps the Hamiltonian to a
quadratic form in fermionic operators, and diagonalization
reduces to a unitary rotation in the single-particle
Hilbert space.  This rotation can be implemented as a
circuit of $\mathcal{O}(n^2)$ Givens rotations, each
costing $\mathcal{O}(1)$ elementary gates, giving a total
diagonalization cost of $\mathcal{O}(n^2)$---polynomial
and practical for moderate~$n$.

\subsection{Connection to Quantum Phase Estimation}
\label{subsec:qpe-connection}

Quantum phase estimation (QPE) can, in principle, extract the
eigenvalues of a Hamiltonian into a register, enabling conditioned
operations on each eigenspace.  This suggests a hybrid approach:
use QPE to resolve eigenspaces, then apply stereographic encoding
conditioned on the QPE register.  However, this reintroduces the
resources that stereographic encoding aims to avoid: QPE requires
$\mathcal{O}(\log(1/\epsilon))$ ancilla qubits and
$\mathcal{O}(\alpha/\epsilon)$ queries to a block encoding
of~$H/\alpha$.  The advantage of stereographic encoding over
standard QSVT in this hybrid setting is unclear and deserves
further investigation.

A concrete hybrid protocol would proceed as follows:
(i)~run QPE with $t$ bits of precision, projecting the
system into an approximate eigenspace;
(ii)~read the QPE register to obtain an estimate
$\tilde{\lambda}$ of the eigenvalue;
(iii)~apply a single-qubit stereographic QSP sequence
on the ancilla using $\tilde{\lambda}$ as the signal
parameter, producing an unbounded rational transformation
$f(\tilde{\lambda})$.  Step~(iii) requires no block encoding
and has depth $\mathcal{O}(d)$ independent of $\|H\|$.
However, steps~(i--ii) require $\mathcal{O}(t)$ controlled
applications of $e^{iH\tau}$ and $t$ ancilla qubits for the
phase register.  The total cost is therefore dominated by
the QPE step: $\mathcal{O}(t \cdot C_{e^{iH\tau}})$, where
$C_{e^{iH\tau}}$ is the cost of implementing the time
evolution.  This hybrid is advantageous only if $C_{e^{iH\tau}}
	\ll \alpha \cdot C_{\mathrm{BE}}$ (the cost of one
$\alpha$-normalized block encoding query), which can occur
for Hamiltonians with efficient Trotter decompositions but
expensive block encodings.

\subsection{Generalized Signal Operators}
\label{subsec:generalized}

Replacing the phase rotations $\exp[i\phi\PauliZ]$ in the QSP
product with general signal-type operators $O_\alpha$
parameterized by $\alpha \in \hat{\C}$ yields a generalized
normalization condition
\begin{equation}
	|P|^2 + |Q|^2
	= (1{+}r^2)^d \textstyle\prod_j(1{+}|\alpha_j|^2),
	\label{eq:generalized-norm}
\end{equation}
whose forward direction follows from the same inductive argument
as the unnormalized formulation
(Section~\ref{subsec:review-function-class}), but whose
converse---decomposing arbitrary polynomial pairs into such a
product---remains open.  Recent work establishing that GQSP is
equivalent to the nonlinear Fourier transform on
$SU(2)$~\cite{Laneve2025b,AlexisBeckerEtAl2026} provides a
natural framework for attacking this question: the
stereographic generalization corresponds to a deformation of the
standard NLFT, and the inverse NLFT algorithms of
Ni et al.~\cite{NiSarkarYingLin2025} could potentially be
adapted to compute phase factors for the generalized product.
This generalization could enlarge the
class of achievable functions and relax the parity constraint.

\subsection{Comparison with Related Work}
\label{subsec:related}

Stereographic block encoding is complementary to several recent
extensions of the QSP framework, each addressing a different
limitation of the standard approach.  We refer the reader to
Lin~\cite{Lin2025} and Joven et al.~\cite{JovenEtAl2025} for
comprehensive surveys of the rapidly evolving landscape.

\emph{Lifting parity and realness.}
GQSP~\cite{MotlaghWiebe2023} removes the parity and realness
constraints of standard QSP and provides a fast angle-finding
algorithm for polynomials of degree~$\sim 10^7$, but the output
remains a bounded polynomial of the block-encoded signal.
Berry et al.~\cite{BerryMotlaghPantaleoniWiebe2024} show that
GQSP can double the efficiency of Hamiltonian simulation by
exploiting controlled forward/reverse block encoding steps.
Modular QSP~\cite{RossiCeroniChuang2025} extends this to
composable multi-input-output gadgets in several variables, and
adversary bounds have been used to characterize achievable
multivariate QSP protocols~\cite{Laneve2025}.  The mathematical
foundations have been deepened through the $SU(2n)$-valued
nonlinear Fourier transform~\cite{AlexisBeckerEtAl2026},
the proof that GQSP and the NLFT on $SU(2)$ are
equivalent~\cite{Laneve2025b}, a numerically stable
inverse NLFT algorithm for phase factor
computation~\cite{NiSarkarYingLin2025}, and a Solovay--Kitaev
theorem for QSP that connects gate approximation theory to
QSP circuit synthesis~\cite{Rossi2025}.
These generalizations enlarge the class of achievable
\emph{bounded} polynomials; stereographic encoding enlarges the
output range instead.

\emph{Eigenvalue transformation beyond QSVT.}
Jiang and An~\cite{JiangAn2026} develop contour-integral based
quantum eigenvalue transformation, which goes beyond the QSVT
framework to implement non-polynomial functions using only three
additional qubits.  Their approach outperforms existing methods
for asymptotically stable ODEs and, like stereographic encoding,
performs eigenvalue (rather than singular value) transformations.
The key difference is that contour-integral methods still require
a standard block encoding as input, inheriting the normalization
overhead $\alpha$, whereas stereographic encoding removes it at
the cost of the eigenbasis requirement.

\emph{Reducing or avoiding block encoding overhead.}
Hamiltonian block encoding~\cite{KangSu2025} achieves quantum
matrix arithmetics with at most two ancilla qubits without
explicit normalization, but still produces bounded
transformations.  Randomized QSVT~\cite{WangZhangHazraEtAl2025}
avoids block encodings entirely through importance sampling, with
gate complexity independent of the number of Hamiltonian terms.
Mahasinghe et al.~\cite{MahasingheEtAl2025} use GQSP with
symmetric combinations of unitary conjugates to synthesize
Hermitian matrix functions without block encoding.  The
power-cosine QSP filter~\cite{Jo2026} achieves ground-state
preparation using only a single ancilla and controlled
time-evolution operators, avoiding block encoding through
mid-circuit measurement and reset.

\emph{Practical block encoding circuits.}
The FOQCS-LCU framework~\cite{NibbiEtAl2026} reduces the
overhead for block-encoding matrix polynomials, with circuit
depth scaling linearly in the polynomial degree and independent
of system size.  Liu et al.~\cite{LiuZhuLowLinYang2025}
achieve $\tilde{\mathcal{O}}(\sqrt{L})$ T-count scaling for
second-quantized Hamiltonians and reduce subnormalization from
$\mathcal{O}(L)$ to $\mathcal{O}(\sqrt{L})$ via
particle-subspace block encoding.
Zecchi et al.~\cite{ZecchiEtAl2026} provide
$\mathcal{O}(\mathrm{poly}(n))$ circuits for banded matrices
with periodic diagonal structure---a class that overlaps with
our diagonal Hamiltonian construction, though their approach
uses LCU rather than eigenbasis-conditioned rotations.
Schillo, Sturm, and Quay~\cite{SchilloSturmQuay2026} introduce
stabilizer-based block encoding of Pauli-string linear
combinations with logarithmic ancilla scaling.
Products between block encodings~\cite{DongLiXue2025} enable
efficient matrix-matrix and Kronecker products with exponential
qubit savings, and the COMPOSER
framework~\cite{PengLiuKowalski2026} provides compile-once
block encodings where geometry updates are absorbed by re-dialed
single-qubit rotations.
Kane et al.~\cite{KaneEtAl2025} use QSVT and
QETU for block encoding lattice boson Hamiltonians, introducing
the LOVE-LCU approach that outperforms asymptotically superior
methods for systems up to $\sim 11$ qubits.  Tang and
Lai~\cite{TangLai2026} extend block encoding to structured
dense matrices via hierarchical low-rank compression.
Concrete resource estimates for block-encoding-based simulation
of scalar field theories~\cite{HardyEtAl2024} indicate that
meaningful calculations require $\sim 4 \times 10^6$ physical
qubits and $\sim 10^{12}$ T-gates, highlighting the importance
of reducing block encoding overhead.

\emph{Exploiting spectral information.}
Suresh~\cite{Suresh2026} shows that when eigenvalues of the
block-encoded matrix are partially known, spectrally corrected
polynomial approximations can reduce QSVT circuit depth by up
to $5\times$.  Deka and Zak~\cite{DekaZak2025} reduce the
block-encoding normalization constant~$\alpha$ by 25\% for
electronic Hamiltonians via combined symmetry shifts and tensor
factorizations, underscoring that the normalization overhead is
a practical bottleneck actively being optimized.
Stereographic encoding takes this to its logical
extreme: when the full eigenbasis is known, the normalization
$\alpha$ can be removed entirely, and the eigenvalue
transformation becomes exact at each eigenspace.

\emph{Noise resilience.}
Liu et al.~\cite{LiuDongAnNiu2026} introduce ensemble-based
QSP, which mitigates coherent phase errors in QSP circuits
through measurement averaging over noisy ensembles, without
increasing circuit depth.  This complements our noise analysis
(Section~\ref{sec:numerics}): the sampling overhead inherent
in stereographic decoding could potentially be combined with
ensemble averaging to suppress both coherent errors and
the additional shot noise from Pauli measurement.  Realistic
resource estimation for early fault-tolerant implementations
can be performed using the FLASQ cost
model~\cite{HugginsEtAl2025}, which captures spacetime
volume overhead including routing and Clifford costs.
Experimentally, programmable bosonic QSP has been demonstrated
in trapped-ion systems for simulating anharmonic
dynamics~\cite{McGarryEtAl2026}, providing the first
experimental validation of QSP-based Hamiltonian simulation.

In summary, stereographic block encoding is the only approach
that produces genuinely unbounded output, at the cost of the
eigenbasis requirement and increased sampling overhead.
It is most naturally compared with contour-integral eigenvalue
transformation~\cite{JiangAn2026}, which also goes beyond
QSVT but retains standard block encoding, and with spectral
correction methods~\cite{Suresh2026}, which exploit partial
spectral knowledge to reduce depth within the standard
framework.  Block-encoding-free
approaches~\cite{MahasingheEtAl2025,
	Jo2026,WangZhangHazraEtAl2025} trade the block encoding
requirement for other constraints.  These approaches target
different points on the information--overhead tradeoff and
are complementary rather than competing.

\subsection{Open Problems}
\label{subsec:open}

Several questions remain open:
\begin{enumerate}
	\item Can the eigenbasis requirement be relaxed for specific
	      Hamiltonian classes, perhaps by exploiting partial
	      spectral information (e.g., knowledge of the ground
	      state but not the full eigenbasis)?
	\item What is the optimal measurement strategy for
	      extracting $f(\lambda_j)$ from a superposition of
	      eigenspaces?  Can adaptive measurement schemes improve
	      the $\mathcal{O}(r^6/\epsilon^2)$ scaling?
	\item Does the generalized normalization
	      condition~\eqref{eq:generalized-norm} admit a
	      completeness theorem analogous to the standard QSP
	      theorem?
	\item Can stereographic encoding be combined with
	      tensor-network techniques for Hamiltonians with
	      known matrix product operator (MPO) structure?
\end{enumerate}
